Methods

### 2021-09-21 Manuscript/Manuscript_Introduction.Rmd
update

Scientific production has experienced a continuous exponential growth in last decades [@larsen2010rate; @bornmann2015growth]. This is especially true for biomedical research, a trend further increased by the COVID19 pandemic, thanks also to a decrease in journals' processing time and to the diffusion of preprint databases' usage [@aviv2021publication; @horbach2020pandemic; @hoy2020rise]. Consequently, it gets harder for researchers and practitioners to stay up-to-date on the latest findings in their field. Secondary research is of paramount relevance in this scenario, providing useful summaries of the state of the art, but is getting ever more demanding in terms of time and human resources [@allen1999estimating; @borah2017analysis; cohen2010evidence; bastian2010seventy].\
The article collection and screening phases of a systematic review are particularly problematic tasks [systematic review tutorial]. The first is the collection of published research possibly relevant to a scientific question through the an appropriately built search queries used with scientific database search engines. The construction of search queries is a particularly delicate tasks [@lefebvre2011searching; @hammerstrom2010searching], requiring both domain and some knowledge of the databases' query languages; the goal is to produce a set of results containing all relevant articles (high sensitivity) and as few non-relevant ones as possible (high specificity), focusing on the first aspect at the cost of the second [@hammerstrom2010searching]. Furthermore, often manual work is required to download, store and organize the publication data. This aspect may be complicated by limits in the number of records that can be downloaded at once [], and by the necessity to harmonize different formats and resolve duplication. The screening phase consists of classifying the collected publications as relevant or not for the topic of interest [@bannach2019machine]. It is usually the more resource demanding task of a systematic review; even with appropriately built search queries, the returned results easily range in the tens of thousands of which just a small fraction is actually relevant [@lefebvre2011searching]. It was estimated that labelling 10 000 publications may take as much as 40 weeks of work, while the average clinical systematic review takes 63 weeks to be completed [@bannach2019machine; @borah2017analysis; @allen1999estimating].\
The field of Data Science applied to evidence synthesis and acquisition has since years reached maturation [@marshall2015systematic; @beller2018making]. Through the application of natural language processing (NLP) it is possible to transform free text into quantitative features, with various level of abstraction and generalization [@ananiadou2006text; @cohen2008getting]; with machine learning, such text-derived data can be used to map and reproduce human judgment to automatize the screening given an initial human labeled set [@ikonomakis2005text].\
The field of automatization of systematic reviews has been ripe of improvements in the last years [@ananiadou2009supporting; @o2015using; @tsafnat2013automation], and it is possible to foresee that such techniques will become a standard approach [@beller2018making]. Some example...
Despite the great advances, each method has a various shortcoming, both in usability and efficiency [@o2015using].
We tried to addresses some of these limitations creating an integrated framework which tries to simplify systematic reviews in innovative ways: our framework provides solutions for automatic publication collection from multiple online sources and management, and a Bayesian, active machine learning based tool to support humans in the screening task, exponentially reducing the number of records to review. Finally, we propose an experimental method to semi-automatically generate efficient search queries given an already labelled set of publications.\
We apply this solution in the production of a systematic review evaluating the mathematical modeling of patient referral networks among hospitals and its impact on the diffusion of healthcare-associated pathogenic microrganisms with a focus on antimicrobial resistant strains. The systematic review protocol is published in [].

>>>>

Scientific production has experienced continuous exponential growth in the last decades [@larsen2010rate; @bornmann2015growth]. This is especially true for biomedical research, a trend further increased by the COVID19 pandemic, thanks to decreased journals' processing time and the diffusion of preprint databases' usage [@aviv2021publication; @horbach2020pandemic; @hoy2020rise]. Consequently, it gets harder for researchers and practitioners to stay up-to-date on the latest findings in their field. Secondary research is of paramount relevance in this scenario, providing valuable summaries of the state of the art, but is getting ever more demanding in terms of time and human resources [@allen1999estimating; @borah2017analysis; @cohen2010evidence; @bastian2010seventy].\
The article collection and screening phases of a systematic review are particularly problematic tasks [systematic review tutorial]. First, relevant published research needs to be collected from scientific databases through appropriately built search queries. The construction of search queries is a particularly delicate task [@lefebvre2011searching; @hammerstrom2010searching], requiring both domain and some knowledge of the databases' query languages; the goal is to produce a set of results containing all relevant articles (high sensitivity) while keeping the total number low (high specificity), focusing on the first aspect at the cost of the second [@hammerstrom2010searching]. If an integrated search tool is not used, manual work is required to download, store and organize the publication data. Searching manually may be complicated by limits in the number of records that can be downloaded at once [] and the necessity to harmonize different formats and resolve duplication. The citation screening phase consists of classifying the collected publications as relevant or not for the topic of interest [@bannach2019machine]. It is usually the more resource-demanding task of a systematic review: even with appropriately built search queries, the returned results easily range in the tens of thousands of which just a small fraction is actually relevant [@lefebvre2011searching]. It was estimated that labelling 10 000 publications may take as much as 40 weeks of work, while the average clinical systematic review takes 63 weeks to be completed [@bannach2019machine; @borah2017analysis; @allen1999estimating]. A consequence is that often systematic reviews are already outdated once they are published [@beller2013systematic].\
The field of Data Science applied to evidence synthesis and acquisition has greatly maturated in the last years [@marshall2015systematic; @beller2018making; @tsafnat2014systematic]. Through the application of natural language processing (NLP), it is possible to transform free text into quantitative features, with various levels of abstraction and generalization [@ananiadou2006text; @cohen2008getting]; with machine learning (ML), such text-derived data can be used to map and reproduce human judgment and automatize the citation screening [@ikonomakis2005text]. One of the most successful approaches to text screening is active ML: researchers review the predictions made by the ML algorithm, focusing on the most uncertain predictions ("uncertainty" based active ML)  or on those giving a document a higher probability of being relevant ("certainty" based active ML) [@miwa2014reducing].\
The automatization of systematic reviews has been ripe with improvements in the last years, and it is possible to foresee that such techniques will become a standard approach [@beller2018making]. See @ananiadou2009supporting; @o2015using; @tsafnat2013automation; @jonnalagadda2015automating for in-depth evaluations of research in this field. Many of these studies have generated a set of user-friendly commercial and free-to-use tools [see @marshall2015systematic, table 1].
Despite the significant advances, each method has various shortcomings, both in usability and efficiency [@o2015using].
We tried to address some of these limitations by creating an integrated framework that tries to simplify systematic reviews in innovative ways: our framework provides solutions for automatic citation collection and management from multiple online sources and a Bayesian, active machine learning-based tool to support humans in the screening task, exponentially reducing the number of records to review. In particular, we exploit the characteristics of Bayesian analysis to overcome the shortcoming of both "certainty" and "uncertainty" based active learning and the difficulties in defining stopping rules [@miwa2014reducing].
Finally, we propose an experimental method to semi-automatically generate efficient search queries given an already labelled set of publications.\
We apply this solution in the production of a systematic review evaluating the mathematical modelling of patient referral networks among hospitals and their impact on the diffusion of healthcare-associated pathogenic microorganisms with a focus on antimicrobial-resistant strains. The systematic review protocol is published in [].

### 2021-09-21 Manuscript/Manuscript_Methods.Rmd
update

### Document-Term matrix generation
The annotation file produced in the previous step has a number of fields characterizing a scientific publication. The framework models the relevance of a record based on the following fields: title, abstract, authors, keywords, MESH terms [@lipscomb2000medical]. A series of Natural Language Processing (NLP) techniques <mark>(*should I put a general reference?*)</mark>) are employed to transform the textual information in these fields into features for machine learning. The processing of free text fields includes tokenization (i.e., extracting the terms), common stopwords (i.e. sentence components bringing no meaning) removal, part-of-speech filtering (only nouns, adjectives, verbs and untagged terms are retained), and lemmatization of the terms (i.e. reduction to their base grammar form). Processing for authors, keywords and MESH terms identify logical units (e.g., author's full names, composite keywords) without breaking them into single terms, and no stopwords removal or lemmatization are applied.\
To reduce noise and save computation time, terms that appeared in less than 5% of the labelled documents (positive and negatives) are removed from negative records. All terms in the positive set are kept to increase sensitivity at the cost of specificity.\
Some terms tend to co-appear in records (non-consecutive ngrams, nc-ngrams), often carrying a particular meaning when copresent. To detect nc-ngrams, we generated a network representation posing edges between terms with a cosine similarity in terms of records copresence > 0.5. We extracted the maximal cliques in the network [@eppstein2010listing] representing highly correlated groups of terms; These generated terms are added to the data set. To avoid overfitting, we kept nc-ngrams of a maximum of ten terms.\
A similarity network is built again using a similarity threshold of 0.9 and finding the cliques again. In this case, the cliques represent terms that always appear together and therefore can be considered synonyms. These terms are merged in the data set to increase computation efficiency and reduce overfitting.\
The final output is a matrix, also called a Document-Term Matrix (DTM), with $N_d$ rows representing the records $D$, $N_t$ terms column for the $t_{field}$ terms (divided by record field) and ${0,1}$ values whether $t_{field} \in D$. We also enriched the DTM with features storing the number of terms in each field to help the model scale term importance based on the document length.

### Label prediction
We used a Bayesian Additive Regression Trees (BART) machine learning model [@chipman2010bart] [in the implementation of @kapelner2013bartmachine] to predict the probability of a record of being relevant, given the information coded into the enriched DTM. BART models have several advantages: as other boosted trees techniques [@hastie2009boosting], they can model complex non-linearities, perform variable selection, manage missing data while sporting high performance in predictive power. However, the Bayesian framework they are built on provides further benefits: less sensitivity on hyperparameter choices, natural regularization, and, most of all, predictive distributions as output in place of point-wise predictions.\
We set up the BART model to use 2000 iterations (after 250 burn-in iterations) and 50 trees; we used a $k$ value of 2 to regularized extreme prediction and let the model use missing fields in the DTM as features [@kapelner2015prediction]. Since the goal is to find all positive matches (i.e., focus on sensitivity), positive records are oversampled ten times.\
The output is a posterior predictive distribution (PPD) of each record's probability of a positive match. An ensemble of ten models was fitted to improve prediction stability by averaging the predictions between models.\
To choose how to label a record, we exploit the uncertainty typical of Bayesian estimates to decide a record label and whether it requires manual evaluation. To describe the process formally, we define

>>>>

### Document-Term matrix generation
The annotation file produced in the previous step has a number of fields characterizing a scientific publication. The framework models the relevance of a record based on the following fields: title, abstract, authors, keywords, MESH terms [@lipscomb2000medical]. A series of Natural Language Processing (NLP) techniques [@marshall2015systematic; @ananiadou2006text] are employed to transform the textual information in these fields into features for machine learning, through a bag-of-words approach [@marshall2015systematic]. The processing of free text fields includes tokenization (i.e., extracting the terms), common stopwords (i.e. sentence components bringing no meaning) removal, part-of-speech filtering (only nouns, adjectives, verbs and untagged terms are retained), and lemmatization of the terms (i.e. reduction to their base grammar form). Processing for authors, keywords and MESH terms identify logical units (e.g., author's full names, composite keywords) without breaking them into single terms, and no stopwords removal or lemmatization are applied.\
To reduce noise and save computation time, terms that appeared in less than 5% of the labelled documents (positive and negatives) are removed from negative records. All terms in the positive set are kept to increase sensitivity at the cost of specificity.\
Some terms tend to co-appear in records (non-consecutive ngrams, nc-ngrams), often carrying a particular meaning when copresent. To detect nc-ngrams, we generated a network representation posing edges between terms with a cosine similarity in terms of records copresence > 0.5. We extracted the maximal cliques in the network [@eppstein2010listing] representing highly correlated groups of terms; These generated terms are added to the data set. To avoid overfitting, we kept nc-ngrams of a maximum of ten terms.\
A similarity network is built again using a similarity threshold of 0.9 and finding the cliques again. In this case, the cliques represent terms that always appear together and therefore can be considered synonyms. These terms are merged in the data set to increase computation efficiency and reduce overfitting.\
The final output is a matrix, also called a Document-Term Matrix (DTM), with $N_d$ rows representing the records $D$, $N_t$ terms column for the $t_{field}$ terms (divided by record field) and ${0,1}$ values whether $t_{field} \in D$. We also enriched the DTM with features storing the number of terms in each field to help the model scale term importance based on the document length.

### Label prediction
We used a Bayesian Additive Regression Trees (BART) machine learning model [@chipman2010bart] [in the implementation of @kapelner2013bartmachine] to predict the probability of a record of being relevant, given the information coded into the enriched DTM. BART models have several advantages: as other boosted trees techniques [@hastie2009boosting], they can model complex non-linearities, perform variable selection, manage missing data while sporting high performance in predictive power. However, the Bayesian framework they are built on provides further benefits: less sensitivity on hyperparameter choices, natural regularization, and, most of all, predictive distributions as output in place of point-wise predictions.\
We set up the BART model to use 2000 iterations (after 250 burn-in iterations) and 50 trees; we used a $k$ value of 2 to regularized extreme prediction and let the model use missing fields in the DTM as features [@kapelner2015prediction]. Since the goal is to find all positive matches (i.e., focus on sensitivity), positive records are oversampled ten times.\
The output is a posterior predictive distribution (PPD) of each record's probability of a positive match. An ensemble of ten models was fitted to improve prediction stability by averaging the predictions between models.\
To choose how to label a record, we exploit the uncertainty typical of Bayesian estimates to decide a record label and whether it requires manual evaluation, an approach joining the "uncertainty" and "certainty" selection criteria of previous work in the field [@miwa2014reducing]. To describe the process formally, we define
